Approximation as Differentiation

### 2025-06-09 main/first-order.tex
Tweaks.

\noindent from which the universal properties of the product and coproduct easily follow. A category $\cat{C}$
with finite biproducts is called \emph{semi-additive}, and cannot be Cartesian closed without being trivial.
Since $X \iso 0 \biprod X$, Cartesian closure implies $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X \Rightarrow
Y} \iso 1$, and all objects are isomorphic.

@@

\noindent Since right adjoints preserve limits and left adjoints preserve colimits, for any $f \adj g: X \to
Y$ we have that $g$ preserves the meet-semilattice structure $(X, \meet, \top)$, and similarly $f$ preserves
the join-semilattice structure $(X, \join, \bot)$. The meet-preserving maps have their $\CMon$ structure
structure (in fact $\SemiLat$ structure, but again $\CMon$ is sufficient): addition of maps is pointwise via
$\meet$, since $(f \meet g)(x) = f(x) \meet g(x)$ preserves meets, with the constant map $\top = x \mapsto
\top_Y$ as a zero. Dually  the join-preserving maps can be added pointwise using $\join$, with the constant
map $\bot = y \mapsto \bot_X$ as a zero. Since these constructions come in adjoint pairs, $\LatGal$ is
enriched in $\CMon$. The hom-set $\LatGal(X,Y)$ provides (component-wise) addition of Galois connections $(f
\adj g) + (f' \adj g') = (f \join f') \adj (g \meet g')$; the Galois connection $\bot \adj \top$ is zero
morphism, and composition is bilinear.

@@

For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{C} = \Set$ and
each category $T(X) = \Func{X}{\cat{C}}$, the functor category which has as objects all $X$-indexed families
of objects of $\cat{C}$ (regarded as functors $F: X \to \cat{C}$) , and as morphisms $F \to G$, all families
of morphisms $\eta(x): F(x) \to G(x)$ in $\cat{C}$ (with naturality trivial because $X$ has only identity
morphisms).

@@

\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus all sets $X$ paired with indexed families
$\partial X: X \to \cat{C}$ and the morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are all
functions $f: X \to Y$ paired with morphisms $\partial f: \partial X \to \partial \reindex{Y}{f}$ in
$\Func{X}{\cat{C}}$.

>>>>

\noindent from which the universal properties of the product and coproduct easily follow. A category $\cat{C}$
with finite biproducts is called \emph{semi-additive}, and cannot be Cartesian closed without being trivial:
since $X \iso 0 \biprod X$, Cartesian closure would imply $\Hom{\cat{C}}{X}{Y} \iso \Hom{\cat{C}}{0}{X
\Rightarrow Y} \iso 1$, and therefore that all objects are isomorphic.

@@

\noindent Since right adjoints preserve limits and left adjoints preserve colimits, for any $f \adj g: X \to
Y$ we have that $g$ preserves the meet-semilattice structure $(X, \meet, \top)$, and similarly $f$ preserves
the join-semilattice structure $(X, \join, \bot)$. The meet-preserving maps are enriched in $\CMon$ (in fact
in $\SemiLat$, but again $\CMon$ is sufficient): addition of maps is computed pointwise via $\meet$, since $(f
\meet g)(x) = f(x) \meet g(x)$ preserves meets, with the constant map $\top = x \mapsto \top_Y$ as a zero.
Dually the join-preserving maps can be added pointwise using $\join$, with the constant map $\bot = y \mapsto
\bot_X$ as a zero. Since these constructions come in adjoint pairs, $\LatGal$ is enriched in $\CMon$. The
hom-set $\LatGal(X,Y)$ provides (component-wise) addition of Galois connections $(f \adj g) + (f' \adj g') =
(f \join f') \adj (g \meet g')$; the Galois connection $\bot \adj \top$ is a zero morphism, and composition is
bilinear.

@@

For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{C} = \Set$ and
each category $T(X) = \Func{X}{\cat{C}}$, the functor category which has as objects all $X$-indexed families
of objects of $\cat{C}$, and as morphisms $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in
$\cat{C}$ (with naturality trivial because $X$ has only identity morphisms).

@@

\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus all pairs of sets $X$ and indexed families
$\partial X: X \to \cat{C}$ and the morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are all
pairs of functions $f: X \to Y$ and morphisms $\partial f: \partial X \to \partial \reindex{Y}{f}$ in
$\Func{X}{\cat{C}}$.
